---
title: "MPAGE: Module-based Pair-wise Analysis of Gene Expression"
author: "Your Name"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MPAGE: Module-based Pair-wise Analysis of Gene Expression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

The **MPAGE** package provides a comprehensive workflow for analyzing RNA modification modules and their dynamic changes across samples. It integrates protein-protein interaction (PPI) network analysis, module identification, and quantification of gene pair relationships to understand the functional organization of RNA modification machinery.

## Installation

You can install MPAGE from GitHub:

```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("loganylchen/MPAGE")
```

```{r setup}
library(mpage)
library(ggplot2)
library(dplyr)
```

## Quick Start Example

Let's walk through a complete analysis using sample data included with the package.

### Step 1: Load Sample Data

```{r load-data}
# Load sample data
data("sample_expression_data")
data("sample_modules")
data("sample_metadata")

# Check dimensions
dim(sample_expression_data)
head(sample_expression_data[, 1:5])
```

### Step 2: Get RNA Modification Proteins

First, we'll retrieve the curated list of RNA modification-related proteins:

```{r get-proteins}
# Get RNA modification proteins
rna_mod_proteins <- get_rna_mod_proteins(
  modification_types = c("m6A", "m5C", "Psi"),
  include_writers = TRUE,
  include_readers = TRUE,
  include_erasers = TRUE
)

head(rna_mod_proteins)
```

### Step 3: Build Protein-Protein Interaction Network

Next, we'll construct a PPI network from the RNA modification proteins:

```{r build-ppi, eval=FALSE}
# Build PPI network (using a small subset for demo)
demo_proteins <- unique(rna_mod_proteins$gene_symbol)[1:20]
ppi_network <- build_ppi_network(
  proteins = demo_proteins,
  data_sources = c("STRING"),
  min_confidence = 0.7
)

# Check network properties
igraph::vcount(ppi_network)
igraph::ecount(ppi_network)
```

Since network construction requires internet access, we'll use the sample modules for demonstration:

```{r use-sample-modules}
# Use sample modules for demonstration
modules <- sample_modules
```

### Step 4: Analyze Module Features

Let's analyze the topological features of our modules:

```{r analyze-features}
# For demonstration, we'll create a mock network
mock_network <- igraph::erdos.renyi.game(50, 0.1)
V(mock_network)$name <- paste0("GENE", 1:50)

# Analyze module features
module_features <- lapply(modules, function(module) {
  subgraph <- igraph::induced_subgraph(mock_network, module$genes)
  list(
    module_id = module$module_id,
    size = length(module$genes),
    density = ifelse(length(module$genes) > 1, igraph::edge_density(subgraph), 0),
    genes = module$genes
  )
})

# Convert to data frame
features_df <- do.call(rbind, lapply(module_features, function(x) {
  data.frame(
    module_id = x$module_id,
    size = x$size,
    density = x$density,
    stringsAsFactors = FALSE
  )
}))

features_df
```

### Step 5: Calculate Module Scores

Now let's calculate module activity scores for our samples:

```{r calculate-scores}
# Use Z-score method for demonstration
module_scores <- score_modules(
  expression_data = sample_expression_data,
  modules = modules,
  methods = c("Zscore")
)

# Check the scores
module_scores$Zscore[1:5, 1:5]
```

### Step 6: Compare Module Scores Across Groups

Let's compare module scores between control and treatment groups:

```{r compare-scores}
# Compare scores across groups
score_comparison <- compare_module_scores(
  module_scores,
  sample_groups = sample_metadata$group,
  method = "Zscore"
)

# Check comparison results
head(score_comparison$comparison_data)
```

### Step 7: Calculate Gene Pair Relationships

Finally, let's analyze gene pair relationships within and between modules:

```{r gene-pairs}
# Calculate intra-module relationships
intra_scores <- calculate_intra_pair_relationships(
  expression_data = sample_expression_data,
  modules = modules,
  metrics = c("correlation")
)

# Check results
names(intra_scores)
length(intra_scores[["m6A_writer"]]$gene_pairs)
```

## Advanced Usage

### Custom RNA Modification Proteins

You can add custom proteins to the curated database:

```{r custom-proteins}
# Define custom proteins
custom_proteins <- data.frame(
  gene_symbol = c("MYGENE1", "MYGENE2"),
  modification_type = c("m6A", "m5C"),
  functional_role = c("writer", "reader"),
  evidence_source = c("MyStudy", "MyStudy")
)

# Add to existing proteins
enhanced_proteins <- add_custom_proteins(rna_mod_proteins, custom_proteins)

# Check new proteins
enhanced_proteins[enhanced_proteins$evidence_source == "MyStudy", ]
```

### Multiple Scoring Methods

MPAGE supports multiple scoring methods for robust analysis:

```{r multiple-methods}
# Note: ssGSEA and GSVA require additional Bioconductor packages
# For demonstration, we'll show the structure
tryCatch({
  multiple_scores <- score_modules(
    expression_data = sample_expression_data,
    modules = modules,
    methods = c("Zscore")
  )
  
  names(multiple_scores)
}, error = function(e) {
  message("Additional packages required for ssGSEA/GSVA")
})
```

## Visualization

MPAGE provides several visualization functions:

```{r visualization}
# Create a simple plot of module scores
df <- data.frame(
  module = rep(sapply(modules, function(x) x$module_id), 20),
  sample = rep(1:20, each = 3),
  score = as.vector(module_scores$Zscore)
)

ggplot(df, aes(x = module, y = score, fill = module)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Module Activity Scores",
       x = "Module",
       y = "Z-score") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Troubleshooting

### Common Issues

1. **Missing dependencies**: Ensure all required packages are installed.
2. **Network construction fails**: Check internet connection for database access.
3. **Low gene overlap**: Verify gene identifiers match between expression data and modules.

### Performance Tips

- Reduce the number of proteins for initial testing
- Use smaller expression datasets for method validation
- Consider parallel processing for large datasets

## Getting Help

- Report issues on [GitHub](https://github.com/loganylchen/MPAGE/issues)
- Check the function documentation with `?function_name`
- Browse the complete workflow in `vignette("MPAGE")`

## Session Info

```{r session-info}
sessionInfo()
```